# Learning

Depending on the [regulatory pathway](https://collaboration.fda.gov/pqigjg5xl6jx/?proto=true), there are several decision points during clinical development that can be used as ["hard" endpoints](https://www.eupati.eu/glossary/hard-endpoint/) to measure "success" by well defined and objective criteria:

(1) Investigational New Drug [(IND)](https://www.fda.gov/drugs/types-applications/investigational-new-drug-ind-application) application or Clinical Trial Application (CTA) granted (yes/no)?  
(3) Market Authorisation granted (yes/no)?  

However, on closer inspection, the criteria to determine (1) and (2) are consdierably "soft" and involve less well-defined terms and subjective judgments such as "substantial evidence" or the ratio of expected benefit and risks. An important task for research ethics is to frame the question of *what determines "translational success"?* in such a way that we devise a metric that is measurable and guarantees that Translational Research froms a learning system that converges towards an optimal trade-off between research-burden-accepte [@lingler2014new, @phillips2019analysis] and information gain [@ioannidis2016most].

Submission, including registration, takes place as part of the [CTA process](https://www.youtube.com/watch?v=9WANVbEgHH8) and will need to include:

- Main characteristics of the trial
- Subject information sheet/Informed Consent Form (ICF)
- Protocol
- Investigational Medicinal Product Dossier(IMPD) (Safety and Efficacy only)
- [Investigator Brochure (IB)]()
- Regulator requests and assessment reports
- Sponsor responses


## Past regulatory applications
With increasing experience on [CTAs](https://eudract.ema.europa.eu/help/Default.htm#eudract/purpose_cta_ov.htm) or [INDs](https://www.fda.gov/drugs/types-applications/investigational-new-drug-ind-application), there may be associations between "success" metrics and certain formal characteristics of application materials. According to [BfArM](https://www.bfarm.de/EN/BfArM/_node.html) there are 3,647 CTA applicatioin for 2014-2018 in Germany and according to EMA there are [132,700 CTA applications](https://eudract.ema.europa.eu/docs/statistics/EudraCT_Statistics_2019/EudraCT_Public_Report_Stats_Oct_2019.pdf) submitted to the EudraCT database in total.

- How many trials fail due to bad methodological [design](https://www.youtube.com/watch?v=hDcedskQy60)? 

```{r eval=TRUE, include=TRUE, results=FALSE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library("dplyr")
library("plotly")
year <- c(2014, 2015, 2016, 2017, 2018)
phase.1 <- c(221, 224, 207, 203, 171)
phase.2 <- c(270, 226, 226, 233, 220)
phase.3 <- c(289, 297, 243, 234, 263)
phase.4 <- c(84, 100, 39, 38, 40)

BfArM.df <- data.frame(matrix(nrow=5, ncol=0)) 
BfArM.df[['Year']] <- year
BfArM.df[['Phase 1']] <- phase.1
BfArM.df[['Phase 2']] <- phase.2
BfArM.df[['Phase 3']] <- phase.3
BfArM.df[['Phase 4']] <- phase.4

BfArM.df



# Get Manufacturer

p <- plot_ly(x = ~year, y = ~phase.1, type = 'bar', name = 'Phase 1') %>%
  add_trace(y = ~phase.2, name = 'Phase 2') %>%
  add_trace(y = ~phase.3, name = 'Phase 3') %>%
  add_trace(y = ~phase.4, name = 'Phase 4') %>%
  layout(yaxis = list(title = 'Count'), barmode = 'stack')

# Create a shareable link to your chart
# Set up API credentials: https://plot.ly/r/getting-started
# chart_link = api_create(p, filename="pie-donut")
# chart_link
```

`r p`

For further information, please refer directly to source of origin [BfArM 2019](https://www.bfarm.de/DE/Arzneimittel/Arzneimittelzulassung/KlinischePruefung/Genehmigungs-Verfahren/Statistik.html).

<!---
![Please refer directly to source of origin. [BfArM 2019](https://www.bfarm.de/DE/Arzneimittel/Arzneimittelzulassung/KlinischePruefung/Genehmigungs-Verfahren/Statistik.html)](https://www.bfarm.de/SharedDocs/Bilder/Arzneimittel/klinPr/Statistik2014-2018.jpg?__blob=normal&v=3){width=450px}


![Please refer directly to source of origin. [Pei 2019](https://www.pei.de/EN/information/license-applicants/clinical-trial-authorisation/statistics/statistics-clinical-trials-node.html)](https://www.pei.de/SharedDocs/Bilder/EN/license-applicants/overview-all-requests.jpg?__blob=thumbnail&v=10){width=450px}
-->

```{r eval=TRUE, include=TRUE, results=FALSE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE }


year <- c(2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018)
phase.1 <- c(10, 24, 27, 27, 36, 30, 33, 47, 36, 41, 48, 51, 55, 73, 53)
phase.2 <- c(4, 43, 59, 89, 92, 79, 97, 113, 118, 96, 83, 125, 131, 117, 151)
phase.3 <- c(10, 74, 86, 83, 73, 97, 108, 107, 100, 100, 119, 143, 141, 130, 136)
phase.4 <- c(1, 14, 10, 11, 12, 13, 9, 13, 11, 8, 17, 8, 13, 12, 9)

Pei.df <- data.frame(matrix(nrow=15, ncol=0)) 
Pei.df[['Year']] <- year
Pei.df[['Phase 1']] <- phase.1
Pei.df[['Phase 2']] <- phase.2
Pei.df[['Phase 3']] <- phase.3
Pei.df[['Phase 4']] <- phase.4

Pei.df

# Get Manufacturer

p2 <- plot_ly(data=Pei.df, x = ~year, y = ~phase.1, type = 'bar', name = 'Phase 1') %>%
  add_trace(y = ~phase.2, name = 'Phase 2') %>%
  add_trace(y = ~phase.3, name = 'Phase 3') %>%
  add_trace(y = ~phase.4, name = 'Phase 4') %>%
  layout(yaxis = list(title = 'Count'), barmode = 'stack')

# Create a shareable link to your chart
# Set up API credentials: https://plot.ly/r/getting-started
# chart_link = api_create(p, filename="pie-donut")
# chart_link
```
`r p2`

For further information, please refer directly to source of origin [Pei 2019](https://www.pei.de/EN/information/license-applicants/clinical-trial-authorisation/statistics/statistics-clinical-trials-node.html).

### Limitations
This kind of data is from [grey literature](https://en.wikipedia.org/wiki/Grey_literature). Although the data is scraped from the web, it can be seen as an "official" release by federal institutions. However, no [API](https://en.wikipedia.org/wiki/Application_programming_interface) is available and the scope and breadth of the data is very limited which severely hampers knowledge from past experience.

[NCATS](https://pubs.acs.org/doi/pdf/10.1021/acsptsci.9b00056)

“Information gain per capita of accepted research burdens” is an useful concept. 
[Intriguing](http://serayamaouche.net/code/R/Rcode.html)


```{r eval=TRUE, include=FALSE, results=FALSE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE }
library("openfda")

# https://github.com/rOpenHealth/openfda
patient_ages = fda_query("/drug/event.json") %>%
               fda_api_key("ex0HLx7faxU5MnmTGJ9GqIRiS2qf0oP6KY8KzF3F") %>%
               fda_count("patient.patientonsetage") %>%
               fda_exec()

## Fetching: https://api.fda.gov/drug/event.json?search=&count=patient.patientonsetage

# patient ages is now a data frame with "term" and "count" columns
# let's plot it with ggplot2
# library("ggplot2")
# qplot(x=term, y=count, data=patient_ages)


SAE_query = fda_query("/drug/event.json") %>%
            fda_api_key("ex0HLx7faxU5MnmTGJ9GqIRiS2qf0oP6KY8KzF3F") %>%
            fda_count("seriousnesscongenitalanomali");

quetia_SEA= SAE_query %>% fda_filter("patient.drug.openfda.generic_name", "quetiapine") %>% fda_exec()

quetia_ages = fda_query("/drug/event.json") %>%
               fda_api_key("ex0HLx7faxU5MnmTGJ9GqIRiS2qf0oP6KY8KzF3F") %>%
               fda_count("patient.patientonsetage") %>%
               fda_exec()


Numb_query = fda_query("/drug/event.json") %>%
            fda_api_key("ex0HLx7faxU5MnmTGJ9GqIRiS2qf0oP6KY8KzF3F") %>%
            fda_count("patient.drug.drugindication");


quetia_num = Numb_query %>% fda_filter("patient.drug.openfda.generic_name", "quetiapine") %>% fda_exec()

```


## EU Open Data Portal
European public assessment reports [(EPAR)](http://data.europa.eu/euodp/en/data/dataset/epar-human-medicines) for human medicines
